Young children rationally integrate semantic and pragmatic information sources to infer the meaning of words

### 2020-09-01 paper/manuscript.Rmd
first draft paper

Experiment 1 tested the so-called mutual exclusivity inference [@markman1988children; @lewis2020-me]. On one table, there was a familiar object, on the other table, there was a novel object (drawn for the purpose of the study) (Fig. 1a/b iv and Fig. S1a). The speaker requested an object by saying "Oh cool, there is a [non-word] on the table, how neat, can you give me the [non-word]?". Children responded by touching one of the objects. The location of the novel object (left or right table) and the animal character were counterbalanced. We coded as correct choice if children chose the novel object as the referent of the novel word. Each child completed 12 trials, each with a different familiar and a different novel object. Familiar objects were selected to vary along the dimension of how likely children were to know the word for it. This including objects that most 2-year-olds can name (e.g. a duck) as well as objects that only very few 5-year-olds can name (e.g. a pawn). The selection was based on age of acquisition ratings from Kuperman and colleagues [-@kuperman2012age]. While these ratings do not capture the absolute age when children acquire these words, they capture the relative order in which words are learned. Fig. S2A in the supplementary material shows the words and objects used in the experiment. We induced this variation to estimate the role of semantic knowledge in a mutual exclusivity inference.

Experiment 2 tested children's sensitivity to common ground that is build up over the course of a conversation. In particular, we tested whether children keep track of which object is new to a speaker and which they have encountered previously [@akhtar1996role; @diesendruck2004two]. The general setup was the same as in Experiment 1 (Fig. S1b). The speaker was positioned between the tables. There was a novel object (drawn for the purpose of the study) on one of the tables while the other table was empty. Next, the speaker turned to one of the tables and either commented on the presence ("Aha, look at that.") or the absence ("Hm, nothing there") of an object. Then the speaker disappeared. While the speaker was away, a second novel object appeared on the previously empty table. Then the speaker returned and requested an object in the same way as in Experiment 1. The positioning of the novel object in the beginning of the experiment, the speaker as well as the location the speaker turned to first was counterbalanced. Children received five trials, each with a different pair of novel objects. We coded as correct choice if children chose the object that was new to the speaker as the referent of the novel word.

Experiment 3 combined the procedures from Experiment 1 and 2. It followed the same procedure as Experiment 2 but involved the same objects as Experiment 1 (Fig. 1 i-iv and Fig. S1c). In the beginning, one table was empty while there was an object (novel or familiar) on the other one. After commenting on the presence or absence of an object on each table, the speaker disappeared and a second object appeared (familiar or novel). Next, the speaker re-appeared and made the usual request. In the congruent condition, the familiar object was present in the beginning and the novel object appeared while the speaker was away (Fig. 1a and Fig. S1c - left). In this case, both the mutual exclusivity and the common ground inference pointed to the novel object as the referent. In the incongruent condition, the novel object was present in the beginning and the familiar object appeared later. In this case, the two inferences pointed to different objects (Fig. 1B and Fig. S1c - right). This resulted in a total of 2 (congruent vs incongurent) x 12 (familiar objects) = 24 different conditions. Participants received up to 12 test trials, six in each condition, each with a different familiar and novel object. Familiar objects were the same as in Experiment 1. The positioning of the objects on the tables, the speaker and the location the speaker first turned to were counterbalanced. Participants could stop the experiment after six trials (three per condition). If a participant stopped after half of the trials, we tested an additional participant to reach a pre-registered number of data points per age group (2-, 3- and 4-year-olds).

@@

To analyze how the manipulations in each experiment affected children's behavior, we used generalized linear mixed models. Since the focus of the paper is on how information sources were integrated, we moved these models to the supplementary material. In the following, we focus on the cognitive models. A detailed, mathematical description, of the different cognitive models along with details about estimation procedures and priors can be found in the supplementary material. All Bayesian cognitive models were implemented in the probabilistic programming language `WebPPL` [@dippl]. The corresponding model code can be found in the associated online repository.

As a first step, we used the data from Experiment 1 and 2 to estimate children's developing sensitivity to each information source. To estimate the parameters for semantic knowledge ($\theta$) and speaker informativeness ($\alpha$), we adapted the rational integration model to a situation in which both objects (novel and familiar) have equal prior probability (i.e., no common ground information). We used the data from Experiment 1 to then infer the parameters. That is, we inferred the intercepts and slopes for speaker informativeness (linear regression) and semantic knowledge (logistic regression) that generated model predictions to match the responses generated in Experiment 1. To estimate the parameters representing sensitivity to common ground ($\rho$), we used a simple logistic regression to infer which combination of intercept and slope would generate predictions that corresponded to the average proportion of correct responses measured in Experiment 2. For the explanation models, these parameters were additionally constrained by the data from Experiment 3. A more detailed description of these parameters were estimated (including a graphical model) can be found in the supplementary material.

To generate model predictions, we combined the parameters according to the respective model formula. As mentioned above, common ground information could either be aligned or in conflict with the other information sources. In the congruent condition, the novel object was also the discourse novel object and thus had the prior probability $\rho$. In the incongruent condition, the novel object was the "old" object and thus had the prior probability of $1 - \rho$.

The rational integration model is defined as $P^{int}_{L_1}(r \mid u; \{\rho_i, \alpha_i\, \theta_{ij}\})\propto P_{S_1}(u \mid r; \{\alpha_i, \theta_{ij}\}) \cdot P(r \mid \rho_i)$. The three lesioned models that were used to compare how well the model predicts new data are reduced versions of this model. The no word knowledge model uses the same model architecture: $P^{no\_wk}_{L_1}(r \mid u; \{\rho_i, \alpha_i\, \theta_{i}\})\propto P_{S_1}(u \mid r; \{\alpha_i, \theta_{i}\}) \cdot P(r \mid \rho_i)$ and the only difference lies in the parameter $\theta$. The object specific parameters for semantic knowledge are fitted via a hierarchical regression (mixed effects) model. In this model, there is an overall developmental trajectory for semantic knowledge (main effect - $\theta_{i}$) and then there is object-specific variation around this trajectory (random effects -- $\theta_{ij}$). Thus, the no word knowledge model takes in the overall trajectory for semantic knowledge ($\theta_{i}$) but ignores object-specific variation. The no common ground model ignores common ground information (represented by $\rho$) and is thus defined as $P^{no\_cg}_{L_1}(r \mid u; \{\alpha_i\, \theta_{ij}\})\propto P_{S_1}(u \mid r; \{\alpha_i, \theta_{ij}\})$. For the no speaker informativeness model, the parameter $\alpha = 0$. As a consequence, the likelihood term in the model is 1 and the model therefore reduces to $P^{no\_si}_{L_1}(r \mid u; \{\rho_i\})\propto P(r \mid \rho_i)$.

>>>>

Experiment 1 tested the so-called mutual exclusivity inference [@markman1988children; @lewis2020-me]. On one table, there was a familiar object, on the other table, there was an unfamiliar object (drawn for the purpose of the study) (Fig. 1a/b iv and Fig. S1a). The speaker requested an object by saying "Oh cool, there is a [non-word] on the table, how neat, can you give me the [non-word]?". Children responded by touching one of the objects. The location of the unfamiliar object (left or right table) and the animal character were counterbalanced. We coded as correct choice if children chose the unfamiliar object as the referent of the novel word. Each child completed 12 trials, each with a different familiar and a different unfamiliar object. Familiar objects were selected to vary along the dimension of how likely children were to know the word for it. This included objects that most 2-year-olds can name (e.g. a duck) as well as objects that only very few 5-year-olds can name (e.g. a pawn). The selection was based on age of acquisition ratings from Kuperman and colleagues [-@kuperman2012age]. While these ratings do not capture the absolute age when children acquire these words, they capture the relative order in which words are learned. Fig. S2A in the supplementary material shows the words and objects used in the experiment.

Experiment 2 tested children's sensitivity to common ground that is build up over the course of a conversation. In particular, we tested whether children keep track of which object is new to a speaker and which they have encountered previously [@akhtar1996role; @diesendruck2004two]. The general setup was the same as in Experiment 1 (Fig. S1b). The speaker was positioned between the tables. There was an unfamiliar object (drawn for the purpose of the study) on one of the tables while the other table was empty. Next, the speaker turned to one of the tables and either commented on the presence ("Aha, look at that.") or the absence ("Hm, nothing there") of an object. Then the speaker disappeared. While the speaker was away, a second unfamiliar object appeared on the previously empty table. Then the speaker returned and requested an object in the same way as in Experiment 1. The positioning of the unfamiliar object in the beginning of the experiment, the speaker as well as the location the speaker turned to first was counterbalanced. Children received five trials, each with a different pair of unfamiliar objects. We coded as correct choice if children chose the object that was new to the speaker as the referent of the novel word.

Experiment 3 combined the procedures from Experiment 1 and 2. It followed the same procedure as Experiment 2 but involved the same objects as Experiment 1 (Fig. 1 i-iv and Fig. S1c). In the beginning, one table was empty while there was an object (unfamiliar or familiar) on the other one. After commenting on the presence or absence of an object on each table, the speaker disappeared and a second object appeared (familiar or unfamiliar). Next, the speaker re-appeared and made the usual request. In the congruent condition, the familiar object was present in the beginning and the unfamiliar object appeared while the speaker was away (Fig. 1a and Fig. S1c - left). In this case, both the mutual exclusivity and the common ground inference pointed to the novel object as the referent. In the incongruent condition, the unfamiliar object was present in the beginning and the familiar object appeared later. In this case, the two inferences pointed to different objects (Fig. 1b and Fig. S1c - right). This resulted in a total of 2 (congruent vs incongurent) x 12 (familiar objects) = 24 different conditions. Participants received up to 12 test trials, six in each alignment condition, each with a different familiar and unfamiliar object. Familiar objects were the same as in Experiment 1. The positioning of the objects on the tables, the speaker and the location the speaker first turned to were counterbalanced. Participants could stop the experiment after six trials (three per alignment condition). If a participant stopped after half of the trials, we tested an additional participant to reach the preregistered number of data points per age group (2-, 3- and 4-year-olds).

@@

To analyze how the manipulations in each experiment affected children's behavior, we used generalized linear mixed models. Since the focus of the paper is on how information sources were integrated, we discuss these models in the supplementary material and focus here on the cognitive models instead. A detailed, mathematical description of the different cognitive models along with details about estimation procedures and priors can be found in the supplementary material. All Bayesian cognitive models were implemented in the probabilistic programming language `WebPPL` [@dippl]. The corresponding model code can be found in the associated online repository.

As a first step, we used the data from Experiment 1 and 2 to estimate children's developing sensitivity to each information source. To estimate the parameters for semantic knowledge ($\theta$) and speaker informativeness ($\alpha$), we adapted the rational integration model to a situation in which both objects (novel and familiar) have equal prior probability (i.e., no common ground information). We used the data from Experiment 1 to then infer the parameters. That is, we inferred the intercepts and slopes for speaker informativeness (linear regression) and semantic knowledge (logistic regression) that generated model predictions to match the responses generated in Experiment 1. To estimate the parameters representing sensitivity to common ground ($\rho$), we used a simple logistic regression to infer which combination of intercept and slope would generate predictions that corresponded to the average proportion of correct responses measured in Experiment 2. For the explanation models, these parameters were additionally constrained by the data from Experiment 3. A more detailed description of how these parameters were estimated (including a graphical model) can be found in the supplementary material.

To generate model predictions, we combined the parameters according to the respective model formula. As mentioned above, common ground information could either be aligned or in conflict with the other information sources. In the congruent condition, the unfamiliar object was also new in context and thus had the prior probability $\rho$. In the incongruent condition, the novel object was the "old" object and thus had the prior probability of $1 - \rho$.

The rational integration model is defined as $P^{int}_{L_1}(r \mid u; \{\rho_i, \alpha_i\, \theta_{ij}\})\propto P_{S_1}(u \mid r; \{\alpha_i, \theta_{ij}\}) \cdot P(r \mid \rho_i)$. The three lesioned models that were used to compare how well the model predicts new data are reduced versions of this model. The no word knowledge model uses the same model architecture: $P^{no\_wk}_{L_1}(r \mid u; \{\rho_i, \alpha_i\, \theta_{i}\})\propto P_{S_1}(u \mid r; \{\alpha_i, \theta_{i}\}) \cdot P(r \mid \rho_i)$ and the only difference lies in the parameter $\theta$. The object specific parameters for semantic knowledge are fitted via a hierarchical regression (mixed effects) model. That is, there is an overall developmental trajectory for semantic knowledge (main effect - $\theta_{i}$) and then there is object-specific variation around this trajectory (random effects -- $\theta_{ij}$). Thus, the no word knowledge model takes in the overall trajectory for semantic knowledge ($\theta_{i}$) but ignores object-specific variation. The no common ground model ignores common ground information (represented by $\rho$) and is thus defined as $P^{no\_cg}_{L_1}(r \mid u; \{\alpha_i\, \theta_{ij}\})\propto P_{S_1}(u \mid r; \{\alpha_i, \theta_{ij}\})$. For the no speaker informativeness model, the parameter $\alpha = 0$. As a consequence, the likelihood term in the model is 1 and the model therefore reduces to $P^{no\_si}_{L_1}(r \mid u; \{\rho_i\})\propto P(r \mid \rho_i)$.

### 2020-09-01 paper/supplementary_material.Rmd
first draft reformatted paper
first draft paper

---
title: "Young children rationally integrate semantic and pragmatic information sources to infer the meaning of words"
subtitle: "Supplementary material"
author: "Manuel Bohn, Michael Henry Tessler, Megan Merrick and Michael C. Frank"
output: 
  bookdown::pdf_document2:
    toc: yes 
    number_sections: false    
  bookdown::html_document2:
    toc: yes
    toc_float: true
    fig_caption: yes
    #code_folding: hide
    number_sections: false
bibliography: library.bib
#csl: pnas.csl
header-includes:
  \usepackage{caption}
  \renewcommand{\thetable}{S\arabic{table}} 
  \renewcommand{\thefigure}{S\arabic{figure}}
---

@@

\begin{equation}
P^{no\_me}_{L_1}(r \mid u; \{\rho_i\})\propto P(r \mid \rho_i) 
(\#eq:nome)
\end{equation}

@@

plot_cor_model_pred <- readRDS( "../saves/corr_model_data_pred.rds")%>%
  mutate(model = factor(model, levels = c("integration", "no word knowledge", "no common ground", "no speaker informativeness")))

>>>>

---
title: "Young children rationally integrate semantic and pragmatic information sources to infer the meaning of words"
subtitle: "Supplementary material"
author: "Manuel Bohn, Michael Henry Tessler, Megan Merrick and Michael C. Frank"
output: 
  bookdown::html_document2:
    toc: yes
    toc_float: true
    fig_caption: yes
    #code_folding: hide
    number_sections: false
  bookdown::pdf_document2:
    toc: yes 
    number_sections: false    
bibliography: library.bib
#csl: pnas.csl
header-includes:
  \usepackage{caption}
  \renewcommand{\thetable}{S\arabic{table}} 
  \renewcommand{\thefigure}{S\arabic{figure}}
---

@@

\begin{equation}
P^{no\_si}_{L_1}(r \mid u; \{\rho_i\})\propto P(r \mid \rho_i) 
(\#eq:nome)
\end{equation}

@@

plot_cor_model_pred <- readRDS( "../saves/corr_model_data_pred.rds")%>%
  mutate(model = ifelse(model == "no mutual exclusivity", "no speaker informativeness", model),
    model = factor(model, levels = c("integration", "no word knowledge", "no common ground", "no speaker informativeness")))
